\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-03-26 doc/sw-cfhtls.tex
PlanetHunters numbers, \PH and \GZ macros

Galaxy Zoo, was the first citizen science project in astronomy which was used to
classify galaxy morphologies \citep{Lintott2008}.  From the early results based
on morphological classifications of galaxies, to several new unexpected and
interesting discoveries, such as that of green pea galaxies and Hanny's
Voorwerp, Galaxy Zoo has been able to harness the potential of citizen
scientists.  Since then astronomy and non-astronomy projects have been launched
under the citizen science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I (REF), we describe the design of \sw and how the entire
system functions as a discovery engine.

>>>>

\GZ was the first citizen science project in astronomy, and addressed the
problem of how to  classify large numbers of galaxies by their morphology
\citep{Lintott2008}.  From the early results based on morphological
classifications of galaxies, to several new unexpected and interesting
discoveries, such as that of green pea galaxies and Hanny's Voorwerp, \GZ
has been able to harness the potential of citizen scientists.  Since then
astronomy and non-astronomy projects have been launched under the citizen
science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I (REF), we describe the design of \sw and how the entire
system functions as a discovery engine.

### 2015-03-26 doc/sw-system.tex
PlanetHunters numbers, \PH and \GZ macros

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project PlanetHunters \citep{SchwambEtal2012}. PlanetHunters was designed to
enable the discovery of transiting exoplanets in data taken by the Kepler
satellite; a community of N inspectors from the general public found, after
each undergoing a small amount of training, N new exoplanet candidates by
visual inspection of the Kepler lightcurves that were presented in a custom
web-based classification interface. The older Galaxy Zoo morphological
classification project \citep{LintottEtal2008} has also enabled the discovery
of rare objects, via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,CC,CM,EB,PM,TJ,LW) were active
in an informal Galaxy Zoo gravitational lens search (Verma et al, in preparation),
an experience which led
to the present hypothesis that a systematic online visual lens search could be
successful.

\todo{PJM}{Fill in numbers of volunteers and detections from PlanetHunters.}

>>>>

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project \PH \citep{SchwambEtal2012}. \PH was designed to enable the discovery of
transiting exoplanets in data taken by the Kepler satellite; a community of over
200,000 inspectors from the general public found, after each undergoing a small
amount of training, over 40 new exoplanet candidates, by visual inspection of
the Kepler lightcurves that were presented in a custom web-based classification
interface \citep{WangEtal2013}. The older \GZ morphological classification
project \citep{LintottEtal2008} has also enabled the discovery of rare objects,
via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,CC,CM,EB,PM,TJ,LW) were
active in an informal \GZ gravitational lens search (Verma et al, in
preparation), an experience which led to the present hypothesis that a
systematic online visual lens search could be successful.
